---
title: "KAP Blood Donation EC"
author: "Htet Phyo Aung"
date: "2025-03-26"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      comment="",
                      message=FALSE,
                      warning = FALSE)
```

```{r packages, include=FALSE}
library(rio)
library(mStats)
library(tidyverse)
library(finalfit)
library(knitr)
library(patchwork)
library(scales)
```


```{r data, include=FALSE}
rm(list = ls())
bd <- readRDS("bd_clean_EC.rds")
mStats::codebook(bd)
dim(bd)
```

## Age distribution of the students

mean

percentiles

sd

present with mean(sd) min max median (IQR)

```{r}
mean(bd$age)
fivenum(bd$age)
sd(bd$age)
```

age distribution histogram

```{r}
bd %>% 
  ggplot(aes(age)) +
  geom_histogram(aes(y= ..density..), bins = 10, fill = "cyan")+
  geom_density(alpha=.2, fill="gray", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(18,35))+
  labs(title="Age Distribution of Students",
       x="Age",
       y="")
```

##frequency table for catagorized variables

```{r}
mStats::tab(bd, sex, marital,living,blood_gp, na.rm = T)

bd %>% filter(blood_gp == "Yes") %>% 
  mStats::tab(what_gp, na.rm = T)

```


## Sex 

```{r}

percent_sex <- bd %>%
  count(sex)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_sex %>%
  ggplot()+
  geom_col(aes(sex,percent,fill= sex))+
  geom_text(aes(sex,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Sex of Students",
       x="",
       y="Percent")

  
```

## Current living condition

```{r}
percent_living <- bd %>%
  count(living)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1))

percent_living %>%
  ggplot()+
  geom_col(aes(living,percent,fill= living))+
  geom_text(aes(living,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Current Living Condition of Students",
       x="",
       y="Percent")
```

# marital status
```{r}

percent_marital<- bd %>%
  count(marital)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1))

percent_marital %>%
  ggplot()+
  geom_col(aes(marital,percent,fill= marital))+
   geom_text(aes(marital,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Marital Status",
       x="",
       y="Percent")
  
```







# Do you know your blood Group?

```{r}

percent_blood_gp <- bd %>%
  count(blood_gp)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1))

percent_blood_gp %>%
  ggplot()+
  geom_col(aes(blood_gp,percent,fill= blood_gp))+
   geom_text(aes(blood_gp,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you know your blood Group?",
       x="",
       y="Percent")
  
```


#Blood Group of students

```{r}

percent_what_gp <- bd %>%
  filter(blood_gp == "Yes") %>% 
  count(what_gp)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1))

percent_what_gp %>%
  ggplot()+
  geom_col(aes(what_gp,percent,fill= what_gp))+
  geom_text(aes(what_gp,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Blood Group of students",
       x="Blood Group",
       y="Percent")
  
```


## Knowledge 

#B1 What is the minimum age for blood donation?

```{r}
bd %>% 
  ggplot(aes(B1)) +
  geom_histogram(aes(y= ..density..), bins = 5, fill = "cyan")+
  geom_density(alpha=.2, fill="gray", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(B1)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(14,26))+
  labs(title="What is the minimum age for blood donation?",
       x="Age",
       y="")
```

#B2 How often can a healthy individual donate blood?
```{r}
bd$B2 <- as.factor(bd$B2)
bd$B2 <- factor(bd$B2,
                        levels=c(1,2,3,4,5,6),
                        labels=c("Once a week", "Once a month", 
                                 "Every 4 months","Every 6 months",
                                 "Once a year","Don't Know"))
```


```{r}

percent_B2 <- bd %>%
  count(B2)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/120,1))

percent_B2 %>%
  ggplot()+
  geom_col(aes(B2,percent,fill= B2))+
  geom_text(aes(B2,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(axis.text.x = element_text(ang= 20))+
  theme(legend.position = "FALSE")+
  labs(title="How often can a healthy individual donate blood?",
       x="",
       y="Percent")

```


#B3 What is the minimum hemoglobin level required for blood donation?

```{r}
bd$B3 <- as.factor(bd$B3)
bd$B3 <- factor(bd$B3,
                        levels=c(1,2,3,4),
                        labels=c("10 g/dL", "12 g/dL", 
                                 "14 g/dL","Don't Know"))
```


```{r}

percent_B3 <- bd %>% 
  count(B3)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/118,1))

percent_B3 %>%
  ggplot()+
  geom_col(aes(B3,percent,fill= B3))+
  geom_text(aes(B3,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="What is the minimum Hb level required for blood donation?",
       x="",
       y="Percent")

```


#B4. Which of the following conditions may disqualify someone from donating blood? (Multiple Response)


```{r}

library(ufs)

qualify <- bd %>% 
  multiResponse(c("Low Hb level" ,"Recent travel to Malaria Area","Recent blood donation","Tatto/Piercing last 6 months","Pregnancy"))

qualify %>% 
  select(`Option`,`Percentage of (119) cases`) %>% 
  arrange(desc(`Percentage of (119) cases`))

```


#B5. What are the potential health benefits of blood donation? (Multiple Response)

```{r}

benefits <- bd %>% 
  multiResponse(c("Reduce Iron level"  ,"Improved CVS health" ,"Burn Calories","Others"))

benefits %>% 
  select(`Option`,`Percentage of (119) cases`) %>% 
  arrange(desc(`Percentage of (119) cases`))

```

#B6. Where can you donate blood? (Multiple Response)

```{r}

donate <- bd %>% 
  multiResponse(c("Hospitals"  ,"Blood Banks" ,"Mobile Donation Units"))

donate %>% 
  select(`Option`,`Percentage of (119) cases`) %>% 
  arrange(desc(`Percentage of (119) cases`))

```

#B7. Where did you get the knowledge about blood donation?(Multiple Response)

```{r}

k_source <- bd %>% 
  multiResponse(c("Health Care Personal","Family/Peer","TV/radio","Internet/Social Media","Journal/Magazine" ))

k_source %>% 
  select(`Option`,`Percentage of (119) cases`) %>% 
  arrange(desc(`Percentage of (119) cases`))

```

# Knowledge score
mean
```{r}
mean(bd$K_score)
sd(bd$K_score)
```
percentiles
```{r}
fivenum(bd$K_score)
```

```{r}
bd <- bd %>% 
  mutate(K_gp = case_when(K_score <= 7  ~ "0",
                              K_score > 7 ~ "1"))
```

```{r}
bd$K_gp <- as.factor(bd$K_gp)
bd$K_gp <- factor(bd$K_gp,
                        levels=c(0,1),
                        labels=c("Poor Knowledge", "Good Knowledge"))
```

```{r}
percent_K_gp <- bd %>% 
  count(K_gp)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_K_gp %>%
  ggplot()+
  geom_col(aes(K_gp,percent,fill= K_gp))+
  geom_text(aes(K_gp,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Knowledge Levels of Students",
       x="",
       y="Percent")

```



## Attitude questions
# C1. Blood donation is a socially responsible act

```{r}
percent_C1 <- bd %>% 
  count(C1)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C1 %>%
  ggplot()+
  geom_col(aes(C1,percent,fill= C1))+
  geom_text(aes(C1,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Blood donation is a socially responsible act",
       x="",
       y="Percent")

```


#C2. I am afraid of the pain or discomfort associated with blood donation

```{r}
percent_C2 <- bd %>% 
  count(C2)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C2 %>%
  ggplot()+
  geom_col(aes(C2,percent,fill= C2))+
  geom_text(aes(C2,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I am afraid of the pain or discomfort 
       associated with blood donation",
       x="",
       y="Percent")

```


#C3. I believe blood donation is beneficial to my health

```{r}
percent_C3 <- bd %>% 
  count(C3)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C3 %>%
  ggplot()+
  geom_col(aes(C3,percent,fill= C3))+
  geom_text(aes(C3,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I believe blood donation is beneficial to my health",
       x="",
       y="Percent")

```


#C4. I am concerned about the risk of contracting diseases through blood donation

```{r}
percent_C4 <- bd %>% 
  count(C4)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C4 %>%
  ggplot()+
  geom_col(aes(C4,percent,fill= C4))+
  geom_text(aes(C4,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I am concerned about the risk of 
       contracting diseases through blood donation",
       x="",
       y="Percent")

```


#C5. I feel proud to donate blood because it helps save lives


```{r}
percent_C5 <- bd %>% 
  count(C5)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C5 %>%
  ggplot()+
  geom_col(aes(C5,percent,fill= C5))+
  geom_text(aes(C5,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I feel proud to donate blood because it helps save lives",
       x="",
       y="Percent")

```

#C6. I think blood donation is inconvenient and time-consuming

```{r}
percent_C6 <- bd %>% 
  count(C6)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C6 %>%
  ggplot()+
  geom_col(aes(C6,percent,fill= C6))+
  geom_text(aes(C6,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I think blood donation is inconvenient and time-consuming",
       x="",
       y="Percent")

```

#C7. I would donate blood if my friends or family encouraged me to do so

```{r}
percent_C7 <- bd %>% 
  count(C7)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C7 %>%
  ggplot()+
  geom_col(aes(C7,percent,fill= C7))+
  geom_text(aes(C7,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I would donate blood if my friends or 
       family encouraged me to do so",
       x="",
       y="Percent")

```

# C8. I believe there is a need for more awareness campaigns about blood donation

```{r}
percent_C8 <- bd %>% 
  count(C8)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_C8 %>%
  ggplot()+
  geom_col(aes(C8,percent,fill= C8))+
  geom_text(aes(C8,percent, label = scales::percent(percent/100)),
            vjust = -0.3, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="I believe there is a need for 
       more awareness campaigns about blood donation",
       x="",
       y="Percent")

```

# Attitude score
mean
```{r}
mean(bd$A_score)
sd(bd$A_score)
```
percentiles
```{r}
fivenum(bd$A_score)
```

```{r}
bd <- bd %>% 
  mutate(A_gp = case_when(A_score <= 25  ~ "0",
                              A_score > 25 ~ "1"))
```

```{r}
bd$A_gp <- as.factor(bd$A_gp)
bd$A_gp <- factor(bd$A_gp,
                        levels=c(0,1),
                        labels=c("Poor Attitude", "Good Attitude"))
```

```{r}
percent_A_gp <- bd %>% 
  count(A_gp)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_A_gp %>%
  ggplot()+
  geom_col(aes(A_gp,percent,fill= A_gp))+
  geom_text(aes(A_gp,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Attitude Levels of Students",
       x="",
       y="Percent")

```

##Practice Questions

#D1. Have you ever donated blood?

```{r}
bd %>% mStats::tab(D1,na.rm = T)
```


```{r}
percent_D1 <- bd %>% 
  count(D1)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/120,1),na.rm = T)

percent_D1 %>%
  ggplot()+
  geom_col(aes(D1,percent,fill= D1))+
  geom_text(aes(D1,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Have you ever donated blood?",
       x="",
       y="Percent")

```

#D2. If yes, how many times have you donated blood?

```{r}
bd %>% 
  filter(D1 == "Yes") %>% 
  mStats::tab(D2)
```

#D3. What motivated you to donate blood? (Multiple Response)

```{r}
moti <- bd %>% 
  filter(D1 == "Yes") %>% 
  multiResponse(c("Altruism_desire to help others","Family/peer influence","Incentives(gifts/certificates)","Health benefits","Other"))

moti %>% 
  select(`Option`,`Percentage of (15) cases`) %>% 
  arrange(desc(`Percentage of (15) cases`))
```

#D4. If you have never donated blood, what are the reasons? (Multiple Response)

```{r}
no_why <- bd %>% 
  filter(D1 == "No") %>% 
  multiResponse(c("Fear of needles/pain" ,"Lack of awareness about B/d donation","Concerns about health risks","Inconvenience","Not eligible to donate"))

no_why %>% 
  select(`Option`,`Percentage of (105) cases`) %>% 
  arrange(desc(`Percentage of (105) cases`))
```


#D5. Would you consider donating blood in the future?

```{r}
percent_D5 <- bd %>% 
  count(D5)%>%
  na.omit() %>% 
  mutate(percent=round(n*100/sum(n),1),na.rm = T)

percent_D5 %>%
  ggplot()+
  geom_col(aes(reorder(D5,-percent),percent,fill= D5))+
  geom_text(aes(D5,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Would you consider donating blood in the future?",
       x="",
       y="Percent")

```


#D6. What would encourage you to donate blood? (Multiple Response)

```{r}
encourage <- bd %>% 
  multiResponse(c("More convenient donation locations","Better awareness campaigns","Incentives(gifts or certificates)","Peer/family encouragement","Others()"))

encourage %>% 
  select(`Option`,`Percentage of (120) cases`) %>% 
  arrange(desc(`Percentage of (120) cases`))
```



