---
title: "eHealth Literacy"
author: "Htet Phyo Aung"
date: "2026-02-08"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      comment="",
                      message=FALSE,
                      warning = FALSE)
```


```{r packages, include=FALSE}
library(tidyverse)
library(tidyquant)
library(ggdist)
library(gghalves)
library(ggpie)
library(ggpubr)
library(ggridges)
library(patchwork)
library(viridis)
library(hrbrthemes)
options(scipen = 999)
```

```{r data, include=FALSE}
setwd("c:/docs/house officers")
rm(list=ls())
eh<- readRDS("eh_clean_ttc.rds")
dim(eh)
names(eh)
```

## (A) Socio-demographics
# Age of participants

mean
percentiles

```{r}
mean(eh$age)
fivenum(eh$age)
```
```{r}
eh %>% 
  ggplot(aes(age)) +
  geom_histogram(aes(y= ..density..), bins = 10, fill = "cyan")+
  geom_density(alpha=.2, fill="cyan", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(16,25))+
  labs(title="Age Distribution of Students",
       x="Age",
       y="")
```

# Percentages for catagorical variables

```{r}
mStats::tab(eh,sex,marital,father_edu, mother_edu,father_occu, mother_occu)
```

#Sex

```{r}
dat1 <- eh %>%
  count(sex)%>%
  arrange(desc(sex)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat1$label <- paste(dat1$sex,"\n",round(dat1$percent),"%",sep="")
dat1

ggplot(dat1,
       aes(x = "",
           y = percent,
           fill = sex)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")

```

#marital status

```{r}
dat2 <- eh %>%
  count(marital)%>%
  arrange(desc(marital)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat2$label <- paste(dat2$marital,"\n",round(dat2$percent),"%",sep="")
dat2

ggplot(dat2,
       aes(x = "",
           y = percent,
           fill = marital)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")
```



#father education status

```{r}
percent_father_edu <- eh %>%
  count(father_edu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_father_edu %>%
  ggplot()+
  geom_col(aes(father_edu,percent,fill= father_edu))+
  geom_text(aes(father_edu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="Father's Education Status",
       x="",
       y="Percent")
```

#mother education status

```{r}
percent_mother_edu <- eh %>%
  count(mother_edu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_mother_edu %>%
  ggplot()+
  geom_col(aes(mother_edu,percent,fill= mother_edu))+
  geom_text(aes(mother_edu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="Mother's Education Status",
       x="",
       y="Percent")
```


#father's Occupation

```{r}
percent_father_occu <- eh %>%
  count(father_occu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_father_occu %>%
  ggplot()+
  geom_col(aes(father_occu,percent,fill= father_occu))+
  geom_text(aes(father_occu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 15))+
  labs(title="Father's Occupation Status",
       x="",
       y="Percent")
```

# mother occupation status

```{r}
percent_mother_occu <- eh %>%
  count(mother_occu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_mother_occu %>%
  ggplot()+
  geom_col(aes(mother_occu,percent,fill= mother_occu))+
  geom_text(aes(mother_occu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Mother's Occupation Status",
       x="",
       y="Percent")
```

# family income per month

mean
percentiles

```{r}
mean(eh$income)
fivenum(eh$income)
```


```{r}
eh %>% ggplot(aes(income)) +
  geom_density(fill= "cyan")+
  theme_minimal()+
  labs(title="Family income per month",
       x="Kyats",
       y="Density")
```


## (B) Internet usage

#B1. Devices for using internet (Multiple response)

```{r}
library(ufs)

eh %>% multiResponse(c("Smart Phone", "Tablet", "Laptop","Desktop","Other"))
```


#B2 Do you have regular access to the internet?

```{r}
dat4 <- eh %>%
  count(B2)%>%
  arrange(desc(B2)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat4$label <- paste(dat4$B2,"\n",round(dat4$percent),"%",sep="")
dat4

ggplot(dat4,
       aes(x = "",
           y = percent,
           fill = B2)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")
```

# B3. average hours per day using the internet?

mean
percentiles

```{r}
mean(eh$B3)
fivenum(eh$B3)
```

```{r}
eh %>% ggplot(aes(B3)) +
  geom_density(fill= "cyan")+
  theme_minimal()+
  labs(title="Average Internet Usage per Day",
       x="Hours",
       y="Density")
```

#B4. used the internet to search for health-related information

```{r}
dat5 <- eh %>%
  count(B4)%>%
  arrange(desc(B4)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat5$label <- paste(dat5$B4,"\n",round(dat5$percent),"%",sep="")
dat5

ggplot(dat5,
       aes(x = "",
           y = percent,
           fill = B4)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")+
  labs(title="Using the internet to search for health-related information",
       x="",
       y="")
```

# B5. When you look for health information online, which sources do you use most often? 

```{r}
eh %>% multiResponse(c("Search engines","Official Website", "General health website", "Social media", "Online Disscussion Group","Mobile Apps", "Others"))
```

## (C) e-Health literacy questionnaire (eHEALS)

# C1. How useful do you feel the Internet is in helping you in making decisions about your health?

```{r}
percent_C1 <- eh %>%
  count(C1)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C1 %>%
  ggplot()+
  geom_col(aes(C1,percent,fill= C1))+
  geom_text(aes(C1,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="How useful do you feel the Internet is in 
       helping you in making decisions about your health?",
       x="",
       y="Percent")
```

#C2. How important is it for you to be able to access health resources on the Internet?

```{r}
percent_C2 <- eh %>%
  count(C2)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C2 %>%
  ggplot()+
  geom_col(aes(C2,percent,fill= C2))+
  geom_text(aes(C2,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="How important is it for you to be able 
       to access health resources on the Internet?",
       x="",
       y="Percent")
```

# C3.I know what health resources are available on the Internet


```{r}
percent_C3 <- eh %>%
  count(C3)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C3 %>%
  ggplot()+
  geom_col(aes(C3,percent,fill= C3))+
  geom_text(aes(C3,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I know what health resources are available on the Internet",
       x="",
       y="Percent")
```

# C4. I know where to find helpful health resources on the Internet

```{r}
percent_C4 <- eh %>%
  count(C4)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C4 %>%
  ggplot()+
  geom_col(aes(C4,percent,fill= C4))+
  geom_text(aes(C4,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I know where to find helpful health 
       resources on the Internet",
       x="",
       y="Percent")
```

# C5. I know how to find helpful health resources on the Internet

```{r}
percent_C5 <- eh %>%
  count(C5)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C5 %>%
  ggplot()+
  geom_col(aes(C5,percent,fill= C5))+
  geom_text(aes(C5,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I know how to find helpful health 
       resources on the Internet",
       x="",
       y="Percent")
```

# C6. I know how to use the Internet to answer my questions about health


```{r}
percent_C6 <- eh %>%
  count(C6)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C6 %>%
  ggplot()+
  geom_col(aes(C6,percent,fill= C6))+
  geom_text(aes(C6,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I know how to use the Internet to 
       answer my questions about health",
       x="",
       y="Percent")
```

# C7. I know how to use the health information I find on the Internet to help me

```{r}
percent_C7 <- eh %>%
  count(C7)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C7 %>%
  ggplot()+
  geom_col(aes(C7,percent,fill= C7))+
  geom_text(aes(C7,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title=" I know how to use the health information 
       I find on the Internet to help me",
       x="",
       y="Percent")
```

# C8. I have the skills I need to evaluate the health resources I find on the Internet

```{r}
percent_C8 <- eh %>%
  count(C8)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C8 %>%
  ggplot()+
  geom_col(aes(C8,percent,fill= C8))+
  geom_text(aes(C8,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I have the skills I need to 
       evaluate the health resources I find on the Internet",
       x="",
       y="Percent")
```

#C9. I can tell high quality health resources from low quality health resources on the Internet

```{r}
percent_C9 <- eh %>%
  count(C9)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C9 %>%
  ggplot()+
  geom_col(aes(C9,percent,fill= C9))+
  geom_text(aes(C9,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I can tell high quality health resources 
       from low quality health resources on the Internet",
       x="",
       y="Percent")
```

# C10. I feel confident in using information from the Internet to make health decisions

```{r}
percent_C10 <- eh %>%
  count(C10)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_C10 %>%
  ggplot()+
  geom_col(aes(C10,percent,fill= C10))+
  geom_text(aes(C10,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="I feel confident in using information 
       from the Internet to make health decisions",
       x="",
       y="Percent")
```

# eHealth Literacy Score

mean
percentiles
```{r}
mean(eh$eHealth_score)
fivenum(eh$eHealth_score)
```

```{r}
eh %>% ggplot(aes(eHealth_score)) +
  geom_density(fill= "cyan")+
  #coord_flip()+
  theme_minimal()+
  labs(title="eHealth Literacy Score",
       x="Score",
       y="Density")
```


```{r}
eh %>%
  group_by(sex) %>%
  mutate(mean_by_sex = mean(eHealth_score)) %>%
  ungroup() %>%
  ggplot(aes(sex, eHealth_score, colour = sex,
             show.legend = F)) +
  coord_flip() +
  geom_jitter(show.legend = F,
              size = 4,
              alpha = 0.2,
              width = 0.05) +
  stat_summary(fun = mean, geom = "point", size = 8, show.legend = F) +
  geom_hline(aes(yintercept = mean(eHealth_score)),
             colour = "gray70",
             linewidth = 0.9) +
  geom_segment(aes(x = sex, xend = sex,
                   y = mean(eHealth_score), yend = mean_by_sex),
               size = 2, show.legend = F) +
  labs(title = "eHealth Literacy scores by Sex",
       x = "Sex",
       y = "eHealth Literacy Score") +
  theme(legend.position = "none") +
  theme_bw()
```





```{r}
eh <- eh %>% 
  mutate(literacy_gp = case_when(eHealth_score <= 31 ~ "0",
                                  eHealth_score > 31 ~ "1"))

eh$literacy_gp <- as.factor(eh$literacy_gp)
eh$literacy_gp <-factor(eh$literacy_gp,
                    levels=c(0,1),
                    labels=c("Low eHealth Literacy", "High eHealth Literacy"))
```

# eHealth Literacy group

```{r}
dat6 <- eh %>%
  count(literacy_gp)%>%
  arrange(desc(literacy_gp)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat6$label <- paste(dat6$literacy_gp,"\n",round(dat6$percent),"%",sep="")
dat6

ggplot(dat6,
       aes(x = "",
           y = percent,
           fill = literacy_gp)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")+
  labs(title="eHealth Literacy Group",
       x="",
       y="")
```

at generated the plot.
