---
title: "Energy Drinks Survey UCH 2026"
author: "Dr. Htet Phyo Aung"
date: "2026-08-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      comment="",
                      message=FALSE,
                      warning = FALSE)
```

```{r packages, include=FALSE}
library(rio)
library(mStats)
library(tidyverse)
library(finalfit)
library(knitr)
library(patchwork)
library(scales)
```

```{r data, include=FALSE}
rm(list = ls())
edrink <- readRDS("edrink_UCH_2026_clean.rds")
mStats::codebook(edrink)
dim(edrink)
```

## Age distribution of the students

mean

percentiles

sd

present with mean(sd) min max median (IQR)

```{r}
mean(edrink$age)
fivenum(edrink$age)
sd(edrink$age)
```

age distribution histogram

```{r}
edrink %>% 
  ggplot(aes(age)) +
  geom_histogram(aes(y= ..density..), 
                 bins = 10, 
                 fill = "cyan")+
  geom_density(alpha=.2, 
               fill="gray", 
               adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(16,26))+
  labs(title="Age Distribution of Students",
       x="Age",
       y="")
```

##frequency table for catagorized variables

```{r}
mStats::tab(edrink, sex, marital,religion,living)
```

## Sex

```{r}

percent_sex <- edrink %>%
  count(sex)%>%
  mutate(percent=round(n*100/139,1))

percent_sex %>%
  ggplot()+
  geom_col(aes(sex,percent,fill= sex))+
  geom_text(aes(sex,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Sex of Students",
       x="",
       y="Percent")

  
```

## Marital status

```{r}
percent_marital <- edrink %>%
  select(marital) %>% 
  na.omit() %>% 
  count(marital)%>%
  mutate(percent=round(n*100/138,1))

percent_marital %>%
  ggplot()+
  geom_col(aes(marital,percent,fill= marital))+
    geom_text(aes(marital,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Marital Status of Students",
       x="",
       y="Percent")
```

## Religion

```{r}
percent_religion <- edrink %>%
  count(religion)%>%
  mutate(percent=round(n*100/139,1))

percent_religion %>%
  ggplot()+
  geom_col(aes(religion,percent,fill= religion))+
    geom_text(aes(religion,
                  percent, 
                  label = scales::percent(percent/100)),
            vjust = -0.1, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Religion of Students",
       x="",
       y="Percent")
```

## expenditure

```{r}
mean(edrink$expenditure)
fivenum(edrink$expenditure)
sd(edrink$expenditure)
```

```{r}
edrink %>% 
  ggplot(aes(expenditure)) +
  geom_histogram(aes(y= ..density..), 
                 bins = 10, 
                 fill = "cyan")+
  geom_density(alpha=.2, 
               fill="gray", 
               adjust = 1.3)+
  geom_vline(aes(xintercept=mean(expenditure)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  #scale_x_continuous(limits = c(16,26))+
  labs(title="Expenditure of Students",
       x="Expenditure",
       y="Density")
```



## Current living condition

```{r}
percent_living <- edrink %>%
  count(living)%>%
  mutate(percent=round(n*100/139,1))

percent_living %>%
  ggplot()+
  geom_col(aes(living,percent,fill= living))+
  geom_text(aes(living,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Current Living Condition of Students",
       x="",
       y="Percent")
```

## friends drinking edrink

```{r}
mStats::tab(edrink, friends)
```

```{r}
percent_friends <- edrink %>%
  count(friends)%>%
  mutate(percent=round(n*100/139,1))

percent_friends %>%
  ggplot()+
  geom_col(aes(friends,percent,fill= friends))+
  geom_text(aes(friends,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do your friends take energy drinks?",
       x="",
       y="Percent")
```

## family take edrink

```{r}
mStats::tab(edrink, family)
```

```{r}
percent_family <- edrink %>%
  count(family)%>%
  mutate(percent=round(n*100/139,1))

percent_family %>%
  ggplot()+
  geom_col(aes(family,percent,fill= family))+
    geom_text(aes(family,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do your family members take energy drinks?",
       x="",
       y="Percent")
```

### Awareness Questions

#B1, Do you know what an energy drink is?

```{r}
percent_know <- edrink %>%
  count(B1)%>%
  mutate(percent=round(n*100/139,1))

percent_know %>%
  ggplot()+
  geom_col(aes(B1,percent,fill= B1))+
  geom_text(aes(B1,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you know what an energy drink is?",
       x="",
       y="Percent")
```

##B2.Choose the kinds of energy drink you know. (multiple response)

```{r}

library(ufs)

name <- edrink %>% 
  multiResponse(c("Red Bull","Shark","LIPOVITAN-D","M-150",
                  "Burn","Speed","Carabao","Rocker","Other"))

name[order(name$`Percentage of (139) cases`, decreasing = TRUE), ]

```




##B3. Which of these substances do you think are contained in energy drinks?(Multiple response)

```{r}
content <- edrink %>% 
  multiResponse(c("Caffeine","Vitamins","Sugars" ,"Flavouring Agents",
                  "Preservatives","Others"))

content[order(content$`Percentage of (139) cases`, decreasing = TRUE), ]
```

##B4. What effects do you think energy drinks have? (Multiple Response)

```{r}
effects <-edrink %>% 
  multiResponse(c("Increase physical resistance" , "Increase study concentration","provide alertness","Don't Know","B4e"))

effects %>% arrange(desc(`Percentage of (139) cases`))
```

##B5. Caffeine in energy drinks causes the following symptoms.(Multiple Response)

```{r}
cause <- edrink %>% 
  multiResponse(c("Headache","Tremor","Nervousness","Palpitation",
                  "Insomnia","Anxiety", "B5g"))
cause %>% arrange(desc(`Percentage of (139) cases`))
```

##B6. Frequent consumption of Energy drinks is

```{r}
mStats::tab(edrink,B6)
```

```{r}
percent_B6 <- edrink %>%
  count(B6)%>%
  mutate(percent=round(n*100/139,1))

percent_B6 %>%
  ggplot()+
  geom_col(aes(B6,percent,fill= B6))+
    geom_text(aes(B6,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Frequent consumption of Energy drinks is",
       x="",
       y="Percent")
```

##B7. Consumption of High-level sugar-sweetened drinks is

```{r}
percent_B7 <- edrink %>%
  count(B7)%>%
  mutate(percent=round(n*100/139,1))

percent_B7 %>%
  na.omit() %>% 
  ggplot()+
  geom_col(aes(B7,percent,fill= B7))+
    geom_text(aes(B7,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Consumption of High-level sugar-sweetened drinks is",
       x="",
       y="Percent")
```

##B8. Health effects/risks due to consumption of energy drinks are(Multiple response)

```{r}
risk <- edrink %>% 
  multiResponse(c("Hyperacidity","Sleep Disturbance","Tuberculosis",
                  "Type 2 Diabetes","Hypertension","Hyperlipidaemia", "Overweight/Obse","Anaemia"))

risk %>% arrange(desc(`Percentage of (139) cases`))
```

### Attitude towards E drinks

##C1. Energy drink consumption is not harmful for health

```{r}
mStats::tab(edrink, C1)
```

```{r}
percent_C1 <- edrink %>%
  count(C1)%>%
  mutate(percent=round(n*100/139,1))

percent_C1 %>%
  ggplot()+
  geom_col(aes(C1,percent,fill= C1))+
    geom_text(aes(C1,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drink consumption is not harmful for health",
       x="",
       y="Percent")
```

##C2. Energy drinks should be consumed because for the nutrients in it

```{r}
mStats::tab(edrink, C2)
```

```{r}
percent_C2 <- edrink %>%
  count(C2)%>%
  mutate(percent=round(n*100/139,1))

percent_C2 %>%
  ggplot()+
  geom_col(aes(C2,percent,fill= C2))+
    geom_text(aes(C2,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drinks should be consumed 
  because for the nutrients in it",
       x="",
       y="Percent")
```

##C3. Energy drinks are consumed because they are easily available

```{r}
mStats::tab(edrink, C3)
```

```{r}
percent_C3 <- edrink %>%
  count(C3)%>%
  mutate(percent=round(n*100/139,1))

percent_C3 %>%
  ggplot()+
  geom_col(aes(C3,percent,fill= C3))+
    geom_text(aes(C3,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drinks are consumed 
       because they are easily available",
       x="",
       y="Percent")
```

##C4. Advertisements for energy drinks should be banned

```{r}
mStats::tab(edrink, C4)
```

```{r}
percent_C4 <- edrink %>%
  count(C4)%>%
  mutate(percent=round(n*100/139,1))

percent_C4 %>%
  ggplot()+
  geom_col(aes(C4,percent,fill= C4))+
    geom_text(aes(C4,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Advertisements for energy drinks 
       should be banned",
       x="",
       y="Percent")
```

##C5. Energy drinks should be sold with cheap price

```{r}
mStats::tab(edrink, C5)
```

```{r}
percent_C5 <- edrink %>%
  count(C5)%>%
  mutate(percent=round(n*100/139,1))

percent_C5 %>%
  ggplot()+
  geom_col(aes(C5,percent,fill= C5))+
    geom_text(aes(C5,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drinks should be sold with cheap price",
       x="",
       y="Percent")
```

##C6. Water or fresh fruit juice should be drunk instead of energy drinks

```{r}
mStats::tab(edrink, C6)
```

```{r}
percent_C6 <- edrink %>%
  count(C6)%>%
  mutate(percent=round(n*100/139,1))

percent_C6 %>%
  ggplot()+
  geom_col(aes(C6,percent,fill= C6))+
    geom_text(aes(C6,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Water or fresh fruit juice should be drunk 
       instead of energy drinks",
       x="",
       y="Percent")
```

##C7. Ingredients in energy drinks should be known

```{r}
mStats::tab(edrink, C7)
```

```{r}
percent_C7 <- edrink %>%
  count(C7)%>%
  mutate(percent=round(n*100/139,1))

percent_C7 %>%
  ggplot()+
  geom_col(aes(C7,percent,fill= C7))+
    geom_text(aes(C7,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Ingredients in energy drinks should be known",
       x="",
       y="Percent")
```

##C8. Energy drinks consumption can cause Diabetes

```{r}
mStats::tab(edrink, C8)
```

```{r}
percent_C8 <- edrink %>%
  count(C8)%>%
  mutate(percent=round(n*100/139,1))

percent_C8 %>%
  ggplot()+
  geom_col(aes(C8,percent,fill= C8))+
    geom_text(aes(C8,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drinks consumption can cause Diabetes",
       x="",
       y="Percent")
```

##C9. Energy drinks are not appropriate for children

```{r}
mStats::tab(edrink, C9)
```

```{r}
percent_C9 <- edrink %>%
  count(C9)%>%
  mutate(percent=round(n*100/139,1))

percent_C9 %>%
  ggplot()+
  geom_col(aes(C9,percent,fill= C9))+
    geom_text(aes(C9,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Energy drinks are not appropriate for children",
       x="",
       y="Percent")
```

##C10. Frequent consumption of energy drinks can become dependent

```{r}
mStats::tab(edrink, C10)
```

```{r}
percent_C10 <- edrink %>%
  count(C10)%>%
  mutate(percent=round(n*100/139,1))

percent_C10 %>%
  ggplot()+
  geom_col(aes(C10,percent,fill= C10))+
    geom_text(aes(C10,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Frequent consumption of energy drinks 
       can become dependent",
       x="",
       y="Percent")
```

###Practice

##D1. Do you recently consumed energy drinks? (within one month)

```{r}
mStats::tab(edrink, D1)
```

```{r}
percent_D1 <- edrink %>%
  count(D1)%>%
  mutate(percent=round(n*100/139,1))

percent_D1 %>%
  ggplot()+
  geom_col(aes(D1,percent,fill= D1))+
    geom_text(aes(D1,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you recently consumed energy drinks?
       (within one month)",
       x="",
       y="Percent")
```

##D2. What age did you start consume energy drink?

```{r}
mean(edrink$D2, na.rm = T)
fivenum(edrink$D2)
sd(edrink$D2, na.rm = T)
```

```{r}
edrink %>% 
  ggplot(aes(D2)) +
  geom_histogram(aes(y= ..density..), 
                 bins = 13, 
                 fill = "cyan")+
  geom_density(alpha=.4,
               fill="gray",
               adjust = 1.3)+
  geom_vline(aes(xintercept=mean(D2, na.rm =T )), 
             color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(0,25))+
  labs(title="Age that starts drinking Energy Drinks",
       x="Age",
       y="")
```

##D3. How many times did you consume energy drink during the last week?

```{r}
mStats::tab(edrink, D3)
```

```{r}
percent_D3 <- edrink %>%
  select(D3) %>% 
  na.omit() %>% 
  count(D3)%>%
  mutate(percent=round(n*100/126,1))

percent_D3 %>%
  ggplot()+
  geom_col(aes(D3,percent,fill= D3))+
    geom_text(aes(D3,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="How many days did you consume energy drink  
       during the last week?",
       x="",
       y="Percent")
```

#D4. How were you first introduced to Energy drinks?

```{r}
mStats::tab(edrink, D4 ,na.rm = T)
```

```{r}
percent_D4 <- edrink %>%
  select(D4) %>% 
  na.omit() %>% 
  count(D4)%>%
  mutate(percent=round(n*100/129,1))

percent_D4 %>%
  ggplot()+
  geom_col(aes(D4,percent,fill= D4))+
    geom_text(aes(D4,percent, label = scales::percent(percent/100)),
            vjust = -0.15, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="How were you first introduced to Energy drinks?",
       x="",
       y="Percent")
```

##D5. Reasons for consuming energy drinks (multiple response)

```{r}
reason <- edrink %>% 
  multiResponse(c("Sports","To stay awake","Mental Enhancer",
                  "Energy Boost","To study","Weight control"))
reason %>% arrange(desc(`Percentage of (129) cases`))
```

#D6 How many energy drinks do you consume in an average week?

```{r}
mean(edrink$D6, na.rm = T)
fivenum(edrink$D6)
sd(edrink$D6, na.rm = T)
```

```{r}
edrink %>% 
  ggplot(aes(D6)) +
  geom_histogram(aes(y= ..density..), 
                 bins = 9, 
                 fill = "cyan")+
  geom_density(alpha=.4,
               fill="gray",
               adjust = 1.3)+
  geom_vline(aes(xintercept=mean(D6, na.rm =T )), 
             color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(0,10))+
  labs(title="Number of Energy Drinks consumed 
       in an average week",
       x="Number of cans",
       y="")
```


##D7. Do you usually read the food label before buying energy drink?

```{r}
mStats::tab(edrink, D7, na.rm = T)
```

```{r}
percent_D7 <- edrink %>%
  select(D7) %>% 
  na.omit() %>% 
  count(D7)%>%
  mutate(percent=round(n*100/129,1))

percent_D7 %>%
  ggplot()+
  geom_col(aes(D7,percent,fill= D7))+
    geom_text(aes(D7,percent,  label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black",
            size = 3)+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you usually read the food label 
       before buying energy drink?",
       x="",
       y="Percent")
```

