---
title: "Health Promoting Lifestyle"
author: "Htet Phyo Aung"
date: "2026-03-01"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      comment="",
                      message=FALSE,
                      warning = FALSE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(tidyquant)
library(ggdist)
library(gghalves)
library(ggpie)
library(ggpubr)
library(ggridges)
library(patchwork)
library(viridis)
library(hrbrthemes)
options(scipen = 999)
```

```{r data, include=FALSE}
setwd("c:/docs/house officers")
rm(list=ls())
hpl<- readRDS("hpl_clean.rds")
dim(hpl)
names(hpl)
```

## (A) Socio-demographics
# Age of participants

mean
percentiles

```{r}
mean(hpl$age)
fivenum(hpl$age)
```
```{r}
hpl %>% 
  ggplot(aes(age)) +
  geom_histogram(aes(y= ..density..), bins = 13, fill = "cyan")+
  geom_density(alpha=.2, fill="cyan", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(16,25))+
  labs(title="Age Distribution of Students",
       x="Age",
       y="")
```


# Percentages for catagorical variables

```{r}
mStats::tab(hpl,sex,marital,father_edu, mother_edu,father_occu, mother_occu)
```

#Sex

```{r}
dat1 <- hpl %>%
  count(sex)%>%
  arrange(desc(sex)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat1$label <- paste(dat1$sex,"\n",round(dat1$percent),"%",sep="")
dat1

ggplot(dat1,
       aes(x = "",
           y = percent,
           fill = sex)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")

```

#marital status

```{r}
dat2 <- hpl %>%
  count(marital)%>%
  arrange(desc(marital)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat2$label <- paste(dat2$marital,"\n",round(dat2$percent),"%",sep="")
dat2

ggplot(dat2,
       aes(x = "",
           y = percent,
           fill = marital)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")
```

#father education status

```{r}
percent_father_edu <- hpl %>%
  count(father_edu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_father_edu %>%
  ggplot()+
  geom_col(aes(father_edu,percent,fill= father_edu))+
  geom_text(aes(father_edu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="Father's Education Status",
       x="",
       y="Percent")
```

#mother education status

```{r}
percent_mother_edu <- hpl %>%
  count(mother_edu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_mother_edu %>%
  ggplot()+
  geom_col(aes(mother_edu,percent,fill= mother_edu))+
  geom_text(aes(mother_edu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  labs(title="Mother's Education Status",
       x="",
       y="Percent")
```

#father's Occupation

```{r}
percent_father_occu <- hpl %>%
  count(father_occu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_father_occu %>%
  ggplot()+
  geom_col(aes(father_occu,percent,fill= father_occu))+
  geom_text(aes(father_occu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 15))+
  labs(title="Father's Occupation Status",
       x="",
       y="Percent")
```

# mother occupation status

```{r}
percent_mother_occu <- hpl %>%
  count(mother_occu)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_mother_occu %>%
  ggplot()+
  geom_col(aes(mother_occu,percent,fill= mother_occu))+
  geom_text(aes(mother_occu,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Mother's Occupation Status",
       x="",
       y="Percent")
```

# family income per month

mean
percentiles

```{r}
mean(hpl$income)
fivenum(hpl$income)
```

```{r}
hpl %>% 
  ggplot(aes(income)) +
  geom_histogram(aes(y= ..density..), bins = , fill = "cyan")+
  geom_density(alpha=.2, fill="cyan", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(income)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  #scale_x_continuous(limits = c(16,25))+
  labs(title="Monthly Family Income of Students",
       x="Kyat",
       y="")
```

## Health Promoting Lifestyle

#B1. Discuss my problems and concerns with people close to me

```{r}
percent_B1 <- hpl %>%
  count(B1)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B1 %>%
  ggplot()+
  geom_col(aes(B1,percent,fill= B1))+
  geom_text(aes(B1,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Discuss my problems and concerns 
       with people close to me",
       x="",
       y="Percent")
```

#B2. Choose a diet low in fat, saturated fat and cholesterol

```{r}
percent_B2 <- hpl %>%
  count(B2)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B2 %>%
  ggplot()+
  geom_col(aes(B2,percent,fill= B2))+
  geom_text(aes(B2,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Choose a diet low in fat, 
       saturated fat and cholesterol",
       x="",
       y="Percent")
```

#B3. Report any unusual signs or symptoms to a physician or other health professional

```{r}
percent_B3 <- hpl %>%
  count(B3)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B3 %>%
  ggplot()+
  geom_col(aes(B3,percent,fill= B3))+
  geom_text(aes(B3,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Report any unusual signs or symptoms 
       to a physician or other health professional",
       x="",
       y="Percent")
```

#B4. follow a planned exercise program

```{r}
percent_B4 <- hpl %>%
  count(B4)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B4 %>%
  ggplot()+
  geom_col(aes(B4,percent,fill= B4))+
  geom_text(aes(B4,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="follow a planned exercise program",
       x="",
       y="Percent")
```

#B5. get enough sleep

```{r}
percent_B5 <- hpl %>%
  count(B5)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B5 %>%
  ggplot()+
  geom_col(aes(B5,percent,fill= B5))+
  geom_text(aes(B5,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Get enough sleep",
       x="",
       y="Percent")
```

#B6. feel I am growing and changing in positive ways

```{r}
percent_B6 <- hpl %>%
  count(B6)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B6 %>%
  ggplot()+
  geom_col(aes(B6,percent,fill= B6))+
  geom_text(aes(B6,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="feel I am growing and changing in positive ways",
       x="",
       y="Percent")
```

#B7. praise other people easily for their achievements

```{r}
percent_B7 <- hpl %>%
  count(B7)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B7 %>%
  ggplot()+
  geom_col(aes(B7,percent,fill= B7))+
  geom_text(aes(B7,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="praise other people easily for their achievements",
       x="",
       y="Percent")
```

#B8. limit use of sugars and food containing sugars (sweets)

```{r}
percent_B8 <- hpl %>%
  count(B8)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B8 %>%
  ggplot()+
  geom_col(aes(B8,percent,fill= B8))+
  geom_text(aes(B8,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="limit use of sugars and food containing sugars (sweets)",
       x="",
       y="Percent")
```

#B9. read or watch TV programs about improving health

```{r}
percent_B9 <- hpl %>%
  count(B9)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B9 %>%
  ggplot()+
  geom_col(aes(B9,percent,fill= B9))+
  geom_text(aes(B9,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Read or watch TV programs about improving health",
       x="",
       y="Percent")
```

#B10. exercise vigorously for 20 or more minutes at least three times a week 

```{r}
percent_B10 <- hpl %>%
  count(B10)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B10 %>%
  ggplot()+
  geom_col(aes(B10,percent,fill= B10))+
  geom_text(aes(B10,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="exercise vigorously for 20 or
       more minutes at least three times a week",
       x="",
       y="Percent")
```

#B11.take some time for relaxation each day

```{r}
percent_B11 <- hpl %>%
  count(B11)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B11 %>%
  ggplot()+
  geom_col(aes(B11,percent,fill= B11))+
  geom_text(aes(B11,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="take some time for relaxation each day",
       x="",
       y="Percent")
```

#B12. Believe that my life has purpose

```{r}
percent_B12 <- hpl %>%
  count(B12)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B12 %>%
  ggplot()+
  geom_col(aes(B12,percent,fill= B12))+
  geom_text(aes(B12,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Believe that my life has purpose",
       x="",
       y="Percent")
```

#B13. maintain meaningful and fulfilling relationships with others

```{r}
percent_B13 <- hpl %>%
  count(B13)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B13 %>%
  ggplot()+
  geom_col(aes(B13,percent,fill= B13))+
  geom_text(aes(B13,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="maintain meaningful and 
       fulfilling relationships with others",
       x="",
       y="Percent")
```

#B14. eat 6-11 serving of bread, cereal, rice and pasta each day

```{r}
percent_B14 <- hpl %>%
  count(B14)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B14 %>%
  ggplot()+
  geom_col(aes(B14,percent,fill= B14))+
  geom_text(aes(B14,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" eat 6-11 serving of bread, 
       cereal, rice and pasta each day",
       x="",
       y="Percent")
```

#B15. question health professionals in order to understand their instruction

```{r}
percent_B15 <- hpl %>%
  count(B15)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B15 %>%
  ggplot()+
  geom_col(aes(B15,percent,fill= B15))+
  geom_text(aes(B15,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" question health professionals 
       in order to understand their instruction",
       x="",
       y="Percent")
```

#B16. take part in light to moderate physical activity 

```{r}
percent_B16 <- hpl %>%
  count(B16)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B16 %>%
  ggplot()+
  geom_col(aes(B16,percent,fill= B16))+
  geom_text(aes(B16,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" take part in light to moderate physical activity",
       x="",
       y="Percent")
```

#B17. accept those things in my life which I cannot change 

```{r}
percent_B17 <- hpl %>%
  count(B17)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B17 %>%
  ggplot()+
  geom_col(aes(B17,percent,fill= B17))+
  geom_text(aes(B17,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" accept those things in my life 
       which I cannot change ",
       x="",
       y="Percent")
```

#B18. look forward to the future

```{r}
percent_B18 <- hpl %>%
  count(B18)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B18 %>%
  ggplot()+
  geom_col(aes(B18,percent,fill= B18))+
  geom_text(aes(B18,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="look forward to the future ",
       x="",
       y="Percent")
```

#B19. spend time with close friends

```{r}
percent_B19 <- hpl %>%
  count(B19)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B19 %>%
  ggplot()+
  geom_col(aes(B19,percent,fill= B19))+
  geom_text(aes(B19,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="spend time with close friends",
       x="",
       y="Percent")
```

#B20. eat 2-4 servings of fruit each day

```{r}
percent_B20 <- hpl %>%
  count(B20)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B20 %>%
  ggplot()+
  geom_col(aes(B20,percent,fill= B20))+
  geom_text(aes(B20,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="eat 2-4 servings of fruit each day",
       x="",
       y="Percent")
```

#B21. get a second opinion when I question my health care provider's advice

```{r}
percent_B21 <- hpl %>%
  count(B21)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B21 %>%
  ggplot()+
  geom_col(aes(B21,percent,fill= B21))+
  geom_text(aes(B21,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="get a second opinion when 
       I question my health care provider's advice",
       x="",
       y="Percent")
```

#B22. take part in leisure-time (recreational) physical activities 

```{r}
percent_B22 <- hpl %>%
  count(B22)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B22 %>%
  ggplot()+
  geom_col(aes(B22,percent,fill= B22))+
  geom_text(aes(B22,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="take part in leisure-time 
       (recreational) physical activities",
       x="",
       y="Percent")
```

#B23. concentrate on pleasant thoughts at bedtime

```{r}
percent_B23 <- hpl %>%
  count(B23)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B23 %>%
  ggplot()+
  geom_col(aes(B23,percent,fill= B23))+
  geom_text(aes(B23,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="concentrate on pleasant thoughts at bedtime",
       x="",
       y="Percent")
```

#B24. feel content and at peace with myself

```{r}
percent_B24 <- hpl %>%
  count(B24)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B24 %>%
  ggplot()+
  geom_col(aes(B24,percent,fill= B24))+
  geom_text(aes(B24,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="feel content and at peace with myself",
       x="",
       y="Percent")
```

#B25. find it easy to show concern, love and warmth to others

```{r}
percent_B25 <- hpl %>%
  count(B25)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B25 %>%
  ggplot()+
  geom_col(aes(B25,percent,fill= B25))+
  geom_text(aes(B25,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" find it easy to show concern, 
       love and warmth to others",
       x="",
       y="Percent")
```

#B26. eat 3-5 servings of vegetables each day

```{r}
percent_B26 <- hpl %>%
  count(B26)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B26 %>%
  ggplot()+
  geom_col(aes(B26,percent,fill= B26))+
  geom_text(aes(B26,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="eat 3-5 servings of vegetables each day",
       x="",
       y="Percent")
```


#B27. discuss my health concerns with health professionals

```{r}
percent_B27 <- hpl %>%
  count(B27)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B27 %>%
  ggplot()+
  geom_col(aes(B27,percent,fill= B27))+
  geom_text(aes(B27,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="discuss my health concerns with health professionals",
       x="",
       y="Percent")
```

#B28. do stretching exercises at least 3 times per week

```{r}
percent_B28 <- hpl %>%
  count(B28)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B28 %>%
  ggplot()+
  geom_col(aes(B28,percent,fill= B28))+
  geom_text(aes(B28,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="do stretching exercises at least 3 times per week",
       x="",
       y="Percent")
```

#B29. use specific methods to control my stress

```{r}
percent_B29 <- hpl %>%
  count(B29)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B29 %>%
  ggplot()+
  geom_col(aes(B29,percent,fill= B29))+
  geom_text(aes(B29,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="use specific methods to control my stress",
       x="",
       y="Percent")
```

#B30. work toward long term goals in my life

```{r}
percent_B30 <- hpl %>%
  count(B30)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B30 %>%
  ggplot()+
  geom_col(aes(B30,percent,fill= B30))+
  geom_text(aes(B30,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="work toward long term goals in my life",
       x="",
       y="Percent")
```

#B31. touch and am touched by people I care about 


```{r}
percent_B31 <- hpl %>%
  count(B31)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B31 %>%
  ggplot()+
  geom_col(aes(B31,percent,fill= B31))+
  geom_text(aes(B31,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="touch and am touched by people I care about",
       x="",
       y="Percent")
```

#B32. Eat 2-3 servings of milk, yogurt or cheese each day

```{r}
percent_B32 <- hpl %>%
  count(B32)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B32 %>%
  ggplot()+
  geom_col(aes(B32,percent,fill= B32))+
  geom_text(aes(B32,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Eat 2-3 servings of milk, 
       yogurt or cheese each day",
       x="",
       y="Percent")
```

#B33. Inspect my body at least monthly for physical changes/danger signs


```{r}
percent_B33 <- hpl %>%
  count(B33)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B33 %>%
  ggplot()+
  geom_col(aes(B33,percent,fill= B33))+
  geom_text(aes(B33,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Inspect my body at least monthly 
       for physical changes/danger signs",
       x="",
       y="Percent")
```

#B34. Get exercise during usual daily activities 


```{r}
percent_B34 <- hpl %>%
  count(B34)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B34 %>%
  ggplot()+
  geom_col(aes(B34,percent,fill= B34))+
  geom_text(aes(B34,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Get exercise during usual daily activities",
       x="",
       y="Percent")
```

#B35. Balance time between work and play

```{r}
percent_B35 <- hpl %>%
  count(B35)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B35 %>%
  ggplot()+
  geom_col(aes(B35,percent,fill= B35))+
  geom_text(aes(B35,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Balance time between work and play",
       x="",
       y="Percent")
```

#B36. find each day interesting and challenging

```{r}
percent_B36 <- hpl %>%
  count(B36)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B36 %>%
  ggplot()+
  geom_col(aes(B36,percent,fill= B36))+
  geom_text(aes(B36,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="find each day interesting and challenging",
       x="",
       y="Percent")
```

#B37. find ways to meet my needs for intimacy

```{r}
percent_B37 <- hpl %>%
  count(B37)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B37 %>%
  ggplot()+
  geom_col(aes(B37,percent,fill= B37))+
  geom_text(aes(B37,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="find ways to meet my needs for intimacy",
       x="",
       y="Percent")
```

#B38. eat only 2-3 servings for the meat, poultry, fish, dried beans, eggs and nuts each day

```{r}
percent_B38 <- hpl %>%
  count(B38)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B38 %>%
  ggplot()+
  geom_col(aes(B38,percent,fill= B38))+
  geom_text(aes(B38,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="eat only 2-3 servings for the meat,
       poultry, fish, dried beans, 
       eggs and nuts each day",
       x="",
       y="Percent")
```


#39. ask for information from health professional about how to take good care of myself

```{r}
percent_B39 <- hpl %>%
  count(B39)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B39 %>%
  ggplot()+
  geom_col(aes(B39,percent,fill= B39))+
  geom_text(aes(B39,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="ask for information from health professional 
       about how to take good care of myself",
       x="",
       y="Percent")
```

#B40. check my pulse rate when exercising

```{r}
percent_B40 <- hpl %>%
  count(B40)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B40 %>%
  ggplot()+
  geom_col(aes(B40,percent,fill= B40))+
  geom_text(aes(B40,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Check my pulse rate when exercising",
       x="",
       y="Percent")
```

#B41. practice relaxation or meditation for 15-20 minutes daily

```{r}
percent_B41 <- hpl %>%
  count(B41)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B41 %>%
  ggplot()+
  geom_col(aes(B41,percent,fill= B41))+
  geom_text(aes(B41,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="practice relaxation or meditation 
       for 15-20 minutes daily",
       x="",
       y="Percent")
```

#B42. I am aware of what is important to me in life

```{r}
percent_B42 <- hpl %>%
  count(B42)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B42 %>%
  ggplot()+
  geom_col(aes(B42,percent,fill= B42))+
  geom_text(aes(B42,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title=" I am aware of what is important 
       to me in life",
       x="",
       y="Percent")
```

#B43. get support from a network of caring people

```{r}
percent_B43 <- hpl %>%
  count(B43)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B43 %>%
  ggplot()+
  geom_col(aes(B43,percent,fill= B43))+
  geom_text(aes(B43,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Get support from a network of caring people",
       x="",
       y="Percent")
```

#B44. Read labels to identify nutrients, fats and sodium content in packaged food

```{r}
percent_B44 <- hpl %>%
  count(B44)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B44 %>%
  ggplot()+
  geom_col(aes(B44,percent,fill= B44))+
  geom_text(aes(B44,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Read labels to identify nutrients, 
       fats and sodium content in packaged food",
       x="",
       y="Percent")
```

#B45. Attend educational programs on personal health care 

```{r}
percent_B45 <- hpl %>%
  count(B45)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B45 %>%
  ggplot()+
  geom_col(aes(B45,percent,fill= B45))+
  geom_text(aes(B45,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Attend educational programs on 
       personal health care",
       x="",
       y="Percent")
```

#B46. Reach my target heart rate when exercising

```{r}
percent_B46 <- hpl %>%
  count(B46)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B46 %>%
  ggplot()+
  geom_col(aes(B46,percent,fill= B46))+
  geom_text(aes(B46,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Reach my target heart rate when exercising",
       x="",
       y="Percent")
```

#B47. Pace myself to prevent tiredness

```{r}
percent_B47 <- hpl %>%
  count(B47)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B47 %>%
  ggplot()+
  geom_col(aes(B47,percent,fill= B47))+
  geom_text(aes(B47,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Pace myself to prevent tiredness",
       x="",
       y="Percent")
```

#B48. Feel connected with some force greater than myself

```{r}
percent_B48 <- hpl %>%
  count(B48)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B48 %>%
  ggplot()+
  geom_col(aes(B48,percent,fill= B48))+
  geom_text(aes(B48,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Feel connected with some force greater than myself",
       x="",
       y="Percent")
```

#B49. Settle conflicts with others through discussion and compromise

```{r}
percent_B49 <- hpl %>%
  count(B49)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B49 %>%
  ggplot()+
  geom_col(aes(B49,percent,fill= B49))+
  geom_text(aes(B49,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Settle conflicts with others through 
       discussion and compromise",
       x="",
       y="Percent")
```

#B50. Eat breakfast

```{r}
percent_B50 <- hpl %>%
  count(B50)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B50 %>%
  ggplot()+
  geom_col(aes(B50,percent,fill= B50))+
  geom_text(aes(B50,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Eat breakfast",
       x="",
       y="Percent")
```

#B51. Seek guidance or counselling when necessary

```{r}
percent_B51 <- hpl %>%
  count(B51)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B51 %>%
  ggplot()+
  geom_col(aes(B51,percent,fill= B51))+
  geom_text(aes(B51,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Seek guidance or counselling when necessary",
       x="",
       y="Percent")
```

#B52. Expose myself to new experiences and challenges 

```{r}
percent_B52 <- hpl %>%
  count(B52)%>%
  mutate(percent=round(n*100/sum(n),1))

percent_B52 %>%
  ggplot()+
  geom_col(aes(B52,percent,fill= B52))+
  geom_text(aes(B52,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  scale_fill_brewer(palette = "Set2")+
  theme(legend.position = "FALSE")+
  theme(axis.text.x = element_text(ang= 12))+
  labs(title="Expose myself to new experiences and challenges",
       x="",
       y="Percent")
```

# Health Promoting Lifestyle Score

mean
percentile

```{r}
mean(hpl$hpl_score)
fivenum(hpl$hpl_score)
```

```{r}
hpl %>% 
  ggplot(aes(hpl_score)) +
  geom_histogram(aes(y= ..density..), bins = , fill = "cyan")+
  geom_density(alpha=.2, fill="cyan", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(hpl_score)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  labs(title="Health Promoting Lifestyle Score",
       x="Score",
       y="")
```








