---
title: "IGD_UMMG"
author: "Htet Phyo Aung"
date: "2025-06-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      comment="",
                      message=FALSE,
                      warning = FALSE)
```

```{r packages, include=FALSE}
library(rio)
library(mStats)
library(tidyverse)
library(finalfit)
library(knitr)
library(patchwork)
library(tidyquant)
library(ggdist)
library(gghalves)
library(ufs)
```

```{r data, include=FALSE}
rm(list = ls())
igd <- readRDS("hs_igd_ummg_clean.rds")
mStats::codebook(igd)
dim(igd)
```

## Age

mean
percentiles

```{r}
mean(igd$age)
fivenum(igd$age)
sd(igd$age)
```
```{r}
igd %>% 
  ggplot(aes(age)) +
  geom_histogram(aes(y= ..density..), bins = 5, fill = "cyan")+
  geom_density(alpha=.2, fill="gray", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(16,24))+
  labs(title="Age Distribution of Students",
       x="Age",
       y="")
```

##frequency table for catagorized variables

```{r}
mStats::tab(igd, sex, year,living, relationship, play_game)
```

## Sex

```{r}

percent_sex <- igd %>%
  count(sex)%>%
  mutate(percent=round(n*100/134,1))

percent_sex %>%
  ggplot()+
  geom_col(aes(sex,percent,fill= sex))+
  geom_text(aes(sex,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Sex of Students",
       x="",
       y="Percent")


```
```{r}
dat1 <- igd %>%
  count(sex)%>%
  arrange(desc(sex)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat1$label <- paste(dat1$sex,"\n",round(dat1$percent),"%",sep="")
dat1

ggplot(dat1,
       aes(x = "",
           y = percent,
           fill = sex)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")
```


## Academic year

```{r}
percent_year <- igd %>%
  count(year)%>%
  mutate(percent=round(n*100/134,1))

percent_year %>%
  ggplot()+
  geom_col(aes(year,percent,fill= year))+
  geom_text(aes(year,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Academic Year of Students",
       x="",
       y="Percent")
```

## Expenditure per month

mean
percentile
sd

```{r}
mean(igd$expend)
fivenum(igd$expend)
sd(igd$expend)
```





## Relationship status

```{r}
percent_relation <- igd %>%
  count(relationship)%>%
  mutate(percent=round(n*100/134,1))

percent_relation %>%
  ggplot()+
  geom_col(aes(relationship,percent,fill= relationship))+
  geom_text(aes(relationship,percent, 
                label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Relationship status of Students",
       x="",
       y="Percent")
```

## Do you ever paly electronic games?

```{r}
dat2 <- igd %>%
  count(play_game)%>%
  arrange(desc(play_game)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat2$label <- paste(dat2$play_game,"\n",round(dat2$percent),"%",sep="")
dat2

ggplot(dat2,
       aes(x = "",
           y = percent,
           fill = play_game)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")
```

## Gaming platform

```{r}
library(ufs)

igd %>% 
  filter(play_game == "Yes") %>% 
  multiResponse(c("Desktop/Laptop" ,"Console(PS3/4,Nintendo)","Mobile Phone/Tablet"))
```


## Reason for Playing Game

```{r}
library(ufs)

igd %>% 
  filter(play_game == "Yes") %>% 
  multiResponse(c("For fun" ,"To pass time" ,"Achievement",
                "To relief stress"))
```

## gaming time in weekdays(Mon-Fri)

mean
percentile
sd
```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::summ(weekday)
```

## gaming time in weekends(Sat/Sun)

mean
percentile
sd
```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::summ(weekend)
```

## weekdays Vs weekends

```{r}
Aa<- igd %>% 
  filter(play_game == "Yes") %>% 
  ggplot(aes(weekday)) + 
  stat_halfeye(adjust = .5, width = .7, .width = 0, justification = -.2, point_colour = NA, fill = "cyan") + 
  geom_boxplot(width = .2, outlier.shape = NA)+
  theme_minimal()+
  theme(legend.position = "FALSE")

Bb <- igd %>% 
  filter(play_game == "Yes") %>% 
  ggplot(aes(weekend)) + 
  stat_halfeye(adjust = .5, width = .7, .width = 0, justification = -.2, point_colour = NA, fill = "cyan") + 
  geom_boxplot(width = .2, outlier.shape = NA)+
  theme_minimal()+
  theme(legend.position = "FALSE")

((Aa)/(Bb)) +
  plot_annotation(title ="Gaming Time in Weekdays Vs Weekends 
                  (Minutes)",
                  theme = theme(plot.title = element_text(size = 18))) 

```

## Cost for gaming

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::summ(game_cost)
```

```{r}
options(scipen=999)
igd %>% 
  filter(play_game == "Yes" & game_cost < 100000) %>% 
  ggplot(aes(game_cost)) +
  geom_histogram(aes(y= ..density..),bins = 15, fill = "cyan")+
  geom_density(alpha=.2, fill="gray", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(game_cost)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(0,200000))+
  labs(title="Cost for gaming",
       x="Kyat",
       y="")
```

## Play online multiplayer games

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::tab("Online_multiplayer games")
```

```{r}
dat3 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(`Online_multiplayer games`)%>%
  arrange(desc(`Online_multiplayer games`)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat3$label <- paste(dat3$`Online_multiplayer games`,
                    "\n",round(dat3$percent),"%",sep="")
dat3

ggplot(dat3,
       aes(x = "",
           y = percent,
           fill = `Online_multiplayer games`)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")+
  labs(title="Do You play online multiplayer games?")
```

## Online gambling games

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::tab("Online_gambling games")
```
```{r}
dat4 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(`Online_gambling games`)%>%
  arrange(desc(`Online_gambling games`)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat4$label <- paste(dat4$`Online_gambling games`,
                    "\n",round(dat4$percent),"%",sep="")
dat4

ggplot(dat4,
       aes(x = "",
           y = percent,
           fill = `Online_gambling games`)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")+
  labs(title="Do You play online gambling games?")
```

## IGD Questions

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::tab(C1, C2, C3, C4, C5, C6, C7, C8, C9)
```


## C1. Do you feel preoccupied with your gaming behavior? 


```{r}
percent_C1 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C1)%>%
  mutate(percent=round(n*100/97,1))

percent_C1 %>%
  ggplot()+
  geom_col(aes(C1,percent,fill= C1))+
  geom_text(aes(C1,percent, 
                label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you feel preoccupied with your gaming behavior?",
       x="",
       y="Percent")

```


## C2. Do you feel more irritability, anxiety or even sadness when you try to either reduce or stop your gaming activity?

```{r}
percent_C2 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C2)%>%
  mutate(percent=round(n*100/97,1))

percent_C2 %>%
  ggplot()+
  geom_col(aes(C2,percent,fill= C2))+
  geom_text(aes(C2,percent, label = scales::percent(percent/100)),
            vjust = -0.2, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you feel more irritability, anxiety 
  or even sadness when you try to either reduce or 
  stop your gaming activity?",
       x="",
       y="Percent")

```

## C3. Do you feel the need to spend increasing amount of time engaged gaming in order to achieve satisfaction or pleasure?

```{r}
percent_C3 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C3)%>%
  mutate(percent=round(n*100/97,1))

percent_C3 %>%
  ggplot()+
  geom_col(aes(C3,percent,fill= C3))+
  geom_text(aes(C3,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you feel the need to spend increasing 
  amount of time engaged gaming in order to achieve 
  satisfaction or pleasure?",
       x="",
       y="Percent")

```

## C4. Do you systematically fail when trying to control or cease your gaming activity?

```{r}
percent_C4 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C4)%>%
  mutate(percent=round(n*100/97,1))

percent_C4 %>%
  ggplot()+
  geom_col(aes(C4,percent,fill= C4))+
  geom_text(aes(C4,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you systematically fail when trying to control or 
       cease your gaming activity?",
       x="",
       y="Percent")

```

## C5. Have you lost interests in previous hobbies and other entertainment activities as a result of your engagement with the game?

```{r}
percent_C5 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C5)%>%
  mutate(percent=round(n*100/97,1))

percent_C5 %>%
  ggplot()+
  geom_col(aes(C5,percent,fill= C5))+
  geom_text(aes(C5,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Have you lost interests in previous hobbies 
  and other entertainment activities as a result 
  of your engagement with the game?",
       x="",
       y="Percent")

```

## C6. Have you continued your gaming activity despite knowing it was causing problems between you and other people

```{r}
percent_C6 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C6)%>%
  mutate(percent=round(n*100/97,1))

percent_C6 %>%
  ggplot()+
  geom_col(aes(C6,percent,fill= C6))+
  geom_text(aes(C6,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Have you continued your gaming activity 
  despite knowing it was causing problems 
  between you and other people",
       x="",
       y="Percent")

```

## C7. Have you deceived any of your family members, therapists or others because the amount of your gaming activity?

```{r}
percent_C7 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C7)%>%
  mutate(percent=round(n*100/97,1))

percent_C7 %>%
  ggplot()+
  geom_col(aes(C7,percent,fill= C7))+
  geom_text(aes(C7,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Have you deceived any of your family 
  members,therapists or others because the amount of your 
  gaming activity?",
       x="",
       y="Percent")

```

## C8. Do you play in order to temporarily escape or relieve a negative mood (e.g., helplessness, guilt, anxiety)?

```{r}
percent_C8 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C8)%>%
  mutate(percent=round(n*100/97,1))

percent_C8 %>%
  ggplot()+
  geom_col(aes(C8,percent,fill= C8))+
  geom_text(aes(C8,percent, 
                label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Do you play in order to temporarily escape 
       or relieve a negative mood 
       (e.g., helplessness, guilt, anxiety)?",
       x="",
       y="Percent")

```

## C9. Have you jeopardized or lost an important relationship, job or an educational or career opportunity because of your gaming activity?

```{r}
percent_C9 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(C9)%>%
  mutate(percent=round(n*100/97,1))

percent_C9 %>%
  ggplot()+
  geom_col(aes(C9,percent,fill= C9))+
  geom_text(aes(C9,percent, label = scales::percent(percent/100)),
            vjust = -0.5, 
            colour="black")+
  scale_fill_brewer(palette = "Set2")+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  theme(legend.position = "FALSE")+
  labs(title="Have you jeopardized or lost an important relationship,
       job or an educational or career opportunity 
       because of your gaming activity?",
       x="",
       y="Percent")

```

## IGD score

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::summ(igd_score)
```


```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  ggplot(aes(igd_score)) +
  geom_histogram(aes(y= ..density..), bins = 15,fill = "cyan")+
  geom_density(alpha=.2, fill="gray", adjust = 1.3)+
  geom_vline(aes(xintercept=mean(age)), color="darkred",
             linetype="dashed")+
  theme_minimal()+
  scale_x_continuous(limits = c(0,40))+
  labs(title="IGD Score",
       x="IGD Score",
       y="")
```

```{r}
igd %>%
  filter(play_game == "Yes") %>% 
  group_by(sex) %>%
  mutate(mean_by_sex = mean(igd_score)) %>%
  ungroup() %>%
  ggplot(aes(sex, igd_score, colour = sex,
             show.legend = F)) +
  coord_flip() +
  geom_jitter(show.legend = F,
              size = 4,
              alpha = 0.2,
              width = 0.05) +
  stat_summary(fun = mean, geom = "point", size = 8, show.legend = F) +
  geom_hline(aes(yintercept = mean(igd_score)),
             colour = "gray70",
             linewidth = 0.9) +
  geom_segment(aes(x = sex, xend = sex,
                   y = mean(igd_score), yend = mean_by_sex),
               size = 2, show.legend = F) +
  labs(title = "IGD scores by Sex",
       x = "Sex",
       y = "IGD Score") +
  theme(legend.position = "none") +
  theme_bw()
```

```{r}
igd %>%
  filter(play_game == "Yes") %>% 
  group_by(year) %>%
  mutate(mean_by_year = mean(igd_score)) %>%
  ungroup() %>%
  ggplot(aes(year, igd_score, colour = year,
             show.legend = F)) +
  coord_flip() +
  geom_jitter(show.legend = F,
              size = 4,
              alpha = 0.2,
              width = 0.05) +
  stat_summary(fun = mean, geom = "point", size = 8, show.legend = F) +
  geom_hline(aes(yintercept = mean(igd_score)),
             colour = "gray70",
             linewidth = 0.9) +
  geom_segment(aes(x = year, xend = year,
                   y = mean(igd_score), yend = mean_by_year),
               size = 2, show.legend = F) +
  labs(title = "IGD scores by Academic Year",
       x = "Year",
       y = "IGD Score") +
  theme(legend.position = "none") +
  theme_bw()
```

## IGD score Group

```{r}
igd %>% 
  filter(play_game == "Yes") %>% 
  mStats::tab(igd_gp)
```


```{r}
dat5 <- igd %>%
  filter(play_game == "Yes") %>% 
  count(igd_gp)%>%
  arrange(desc(igd_gp)) %>%
  mutate(percent = round(n*100/sum(n),1),
         lab_pos = cumsum(percent)-0.5*percent)

dat5$label <- paste(dat5$igd_gp,
                    "\n",round(dat5$percent),"%",sep="")
dat5

ggplot(dat5,
       aes(x = "",
           y = percent,
           fill = igd_gp)) +
  geom_bar(width = 1,
           stat = "identity",
           color = "black") +
  geom_text(aes(y = lab_pos, label = label),
            color = "black") +
  coord_polar("y",
              start = 0,
              direction = -1) +
  scale_fill_brewer(palette = "Set2")+ 
  theme_void()+
  theme(legend.position = "FALSE")+
  labs(title="IGD score group")
```

```{r}
igd %>%
  filter(play_game == "Yes") %>% 
  group_by(living) %>%
  mutate(mean_by_living = mean(igd_score)) %>%
  ungroup() %>%
  ggplot(aes(living, igd_score, colour = living,
             show.legend = F)) +
  coord_flip() +
  geom_jitter(show.legend = F,
              size = 4,
              alpha = 0.2,
              width = 0.05) +
  stat_summary(fun = mean, geom = "point", size = 8, show.legend = F) +
  geom_hline(aes(yintercept = mean(igd_score)),
             colour = "gray70",
             linewidth = 0.9) +
  geom_segment(aes(x = living, xend = living,
                   y = mean(igd_score), yend = mean_by_living),
               size = 2, show.legend = F) +
  labs(title = "IGD scores by Living status",
       x = "Living Status",
       y = "IGD Score") +
  theme(legend.position = "none") +
  theme_minimal()
```



